---
title: "Getting Started with lbugr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with lbugr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
  - |
    ```{=html}
    <style>
      body { background-color: #0A0A1A; color: #FFFFFF; }
      h1, h2, h3 { color: #FFFFFF; }
      h1 { background: linear-gradient(135deg, #DC2626 0%, #EF4444 50%, #B91C1C 100%); -webkit-background-clip: text; -webkit-text-fill-color: transparent; background-clip: text; }
      a { color: #EF4444; }
      a:hover { color: #DC2626; }
      code { background-color: #111827; color: #94A3B8; border: 1px solid #1F2937; }
      pre { background-color: #111827; color: #94A3B8; border: 1px solid #1F2937; }
      table { border-color: #1F2937; }
      th { background-color: #111827; color: #FFFFFF; border-color: #1F2937; }
      td { border-color: #1F2937; }
      blockquote { border-left-color: #DC2626; background-color: rgba(220, 38, 38, 0.1); }
    </style>
    ```
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
skip_vignette <- !reticulate::py_module_available("ladybug")
```

## Introduction

Welcome to `lbugr`! This guide will walk you through the basic steps to get started with `lbugr`, from installation to running your first query. `lbugr` is the R interface to the Ladybug Graph Database, a fork of the Kuzu graph database.

## Installation

First, ensure you have the `lbugr` package installed. You will also need `reticulate` to manage the Python environment.

```{r, eval=FALSE}
# Install lbugr from GitHub
remotes::install_github("your-github-repo/lbugr")


# Install the ladybug Python package
reticulate::py_install("ladybug", pip = TRUE)
```

## Basic Usage

### 1. Create a Connection

The first step is to create a connection to a Ladybug database. You can create an in-memory database or connect to a database on disk.

```{r, eval=!skip_vignette}
library(lbugr)

# Create an in-memory database connection
con <- lb_connection(":memory:")
```

### 2. Create a Schema

Next, define your graph schema using Cypher queries. Let's create a simple schema with `Person` nodes and `Knows` relationships.

```{r, eval=!skip_vignette}
lb_execute(con, paste("CREATE NODE TABLE Person(name STRING, age INT64,",
                        "PRIMARY KEY (name))"))
lb_execute(con, "CREATE REL TABLE Knows(FROM Person TO Person, since INT64)")
```

### 3. Load Data

You can load data from R data frames directly into your Ladybug database.

```{r, eval=!skip_vignette}
# Create a data frame of persons
persons_df <- data.frame(
  name = c("Alice", "Bob", "Carol"),
  age = c(35, 45, 25)
)

# Create a data frame of relationships
knows_df <- data.frame(
  from_person = c("Alice", "Bob"),
  to_person = c("Bob", "Carol"),
  since = c(2010, 2015)
)

# Load data into Ladybug
lb_copy_from_df(con, persons_df, "Person")
lb_copy_from_df(con, knows_df, "Knows")
```

### 4. Query Data

Finally, you can query your graph using Cypher and retrieve the results as an R data frame.

```{r, eval=!skip_vignette}
# Execute a query
result <- lb_execute(con, paste("MATCH (a:Person)-[k:Knows]->(b:Person)",
                                  "RETURN a.name, b.name, k.since"))

# Convert the result to a data frame
df <- as.data.frame(result)
print(df)
```

This concludes the "Getting Started" guide. For more advanced topics, please see the other articles and the function reference.
